\Large The psychological arrow of time drives temporal asymmetries in inferring unobserved past and future events

### 2023-09-18 paper/main.tex
checking in new (minor) edits

\title{\Large The psychological arrow of time drives temporal asymmetries in inferring unobserved past and future events}
\author[1]{Xinming Xu}
\author[2]{Ziyan Zhu}
\author[3]{Xueyao Zheng}
\author[1, $\star$]{Jeremy R. Manning}
\affil[1]{Dartmouth College, Hanover, NH, USA}
\affil[2]{Peking University, Beijing, China}
\affil[3]{University of California at Davis, CA, USA}
\affil[$\star$]{Address correspondence to jeremy.r.manning@dartmouth.edu}

@@

  \caption{\textbf{Characters' references drive participants' retrodiction and
  prediction performance.} \textbf{A. Illustration of annotation approach.} We
  manually annotated references to events in past or future segments in
  characters' spoken conversations. We matched each such reference with its
  corresponding storyline event (and its corresponding segment number for
  onscreen events, or half-step segment number for offscreen events). We then
  tracked the hit rate separately for referenced versus unreferenced events in
  participants' uncued retrodictions and predictions. \textbf{B. Reference rate
  as a function of lag.} Across all possible just-watched segments (lag 0), the
  bar heights denote the average proportions of events referenced in other past
  (orange, negative lags) or future (blue, positive lags) segments in our main
  experiment's stimulus. \textbf{C. Difference in hit rates between all events
  and unreferenced events.} To highlight the effect of characters' references
  to past and future events on participants' retrodictions and predictions,
  here we display the difference in across-segment mean hit rates between all
  events and unreferenced events, as a function of temporal distance (lag) to
  the just-watched segment. \textbf{D. Hit rates for unreferenced events.} The
  average response hit rates for unreferenced events are displayed as a
  function of temporal distance to the just-watched segment. Error bars denote
  bootstrapped 95\% confidence intervals. Panels B--D: colors are described in
  the Figure~\ref{fig:result2} caption. \textbf{E. Hit rates and counts of
  referenced and unreferenced events.} As a function of temporal distance to
  the just-watched segment, the sub-panels display the across-segment mean
  numbers ($x$-axes) and hit rates ($y$-axes) of referenced (red) and
  unreferenced (gray) events that participants hit (darker shading) or missed
  (lighter shading) in their uncued retrodictions (top sub-panel) and uncued
  predictions (bottom sub-panel). For an analogous presentation of results from
  the replication experiment, see Fig.~\characterRefs.}

@@

  \caption{\textbf{How much information about the past and future can be
  inferred by observing the present?} By definition, let us say that the
  present moment (lag 0) contains all information about itself (dark gray).
  Given learned statistical regularities, one might extrapolate from the
  present moment into the past or future (light gray). As illustrated in this
  schematic, the information contained in the present about other moments in
  time falls off with absolute lag. This falloff is approximately
  time-symmetric. References in the present to past events (dark orange) or
  future events (dark blue) provide additional information about those
  referenced moments in time, beyond what could be inferred solely from
  statistical regularities. This additional information about those referenced
  moments can also be extrapolated to other moments that are temporally nearby
  to \textit{them} (light orange and blue).}

>>>>

\title{\Large The psychological arrow of time drives temporal asymmetries in inferring unobserved past and future events}
\author[1]{Xinming Xu}
\author[2]{Ziyan Zhu}
\author[3]{Xueyao Zheng}
\author[1, $\star$]{Jeremy R. Manning}
\affil[1]{Dartmouth College, Hanover, NH, USA}
\affil[2]{Peking University, Beijing, China}
\affil[3]{Beijing Normal University, Beijing, China}
\affil[$\star$]{Address correspondence to jeremy.r.manning@dartmouth.edu}

@@

  \caption{\textbf{Characters' references drive participants' retrodiction and
  prediction performance.} \textbf{A. Illustration of annotation approach.} We
  manually annotated references to events in past or future segments in
  characters' spoken conversations. We matched each such reference with its
  corresponding storyline event (and its corresponding segment number for
  onscreen events, or half-step segment number for offscreen events). We then
  tracked the hit rate separately for referenced versus unreferenced events in
  participants' uncued retrodictions and predictions. \textbf{B. Reference rate
  as a function of lag.} Across all possible just-watched segments (lag 0), the
  bar heights denote the average proportions of events referenced in other past
  (orange, negative lags) or future (blue, positive lags) segments in our main
  experiment's stimulus. \textbf{C. Difference in hit rates between all events
  and unreferenced events.} To highlight the effect of characters' references
  to past and future events on participants' retrodictions and predictions,
  here we display the difference in across-segment mean hit rates between all
  events and unreferenced events, as a function of temporal distance (lag) to
  the just-watched segment. \textbf{D. Hit rates for unreferenced events.} The
  average response hit rates for unreferenced events are displayed as a
  function of temporal distance to the just-watched segment. Error bars denote
  bootstrapped 95\% confidence intervals. Panels B--D: colors are described in
  the Figure~\ref{fig:result2} caption. \textbf{E. Hit rates and counts of
  referenced and unreferenced events.} As a function of temporal distance to
  the just-watched segment, the sub-panels display the across-segment mean
  numbers ($x$-axes) and hit rates ($y$-axes) of referenced (red) and
  unreferenced (gray) events that participants hit (darker shading) or missed
  (lighter shading) in their uncued retrodictions (top sub-panel) and uncued
  predictions (bottom sub-panel). Intuitively, the widths of the rectangles at
  each lag denote the total number of events at each possible lag. The darker
  shading denotes the proportions of events that participants retrodicted or
  predicted, and the lighter shading denotes the proportions of events that
  participants ``missed'' in their responses. For an analogous presentation of
  results from the replication experiment, see Fig.~\characterRefs.}

@@

  \caption{\textbf{How much information about the past and future can be
  inferred by observing the present?} By definition, let us say that the
  present moment (lag 0) contains all information about itself (dark gray).
  Given learned statistical regularities, one might extrapolate from the
  present moment into the past or future (light gray). As illustrated in this
  schematic, the information contained in the present about other moments in
  time falls off with absolute lag. This falloff is approximately
  time-symmetric. References in the present to past events (dark orange) or
  future events (dark blue) provide additional information about those
  referenced moments in time, beyond what could be inferred solely from
  statistical regularities. This additional information about those referenced
  moments can also be extrapolated to other moments that are temporally nearby
  to \textit{them} (light orange and blue). The data in this schematic are
  hypothetical.}
